Project Progress Report

### 2017-03-15 pnsac/newsletter/article_MemberInterview.tex
Draft 3--sent to Roger for review.

In 1968 I joined Air Canada as a junior engineer in the Mechanical
Systems Department and assigned as a landing gear technologist until I
achieved Professional Engineer status in 1974. I then moved around
several other departments in engineering and maintenance until I was
promoted to Superintendent of DC9 maintenance. In that capacity I was
involved in the accident investigation into the loss of one aircraft at
Cincinnati.

@@

I worked on the Project for 10 years starting in 2003 both as a “hands
on” volunteer as well as a member of the board of directors and officer
of the Association. Generally I put in at least two days per week and
sometimes more.

>>>>

In 1968 I joined Air Canada as a junior engineer in the Mechanical
Systems Department and assigned as a landing gear technologist until I
achieved Professional Engineer status in 1974. I then moved around
several other departments in engineering and maintenance until I was
promoted to Superintendent of DC 9 maintenance. In that capacity I was
involved in the accident investigation into the loss of one aircraft at
Cincinnati.

@@

I worked on the Project for 10 years starting in 2003 both as a "hands
on" volunteer as well as a member of the board of directors and officer
of the Association. Generally I put in at least two days per week and
sometimes more.

### 2017-03-15 pnsac/newsletter/article_PresReport.tex
Draft 3--sent to Roger for review.

The last North Star Chronicle was published in March 2016 and since that time we
have successfully continued our work on the aircraft while the Museum was making
preparations for the Star Trek event in the Reserve Hangar. We were not asked to
suspend our work, unlike in 2014 when preparations for the Star Wars event were
being made. This has helped us to retain our Restoration volunteers and recruit
two new ones. However, we are still short of Restoration volunteers and continue
to actively work with the Museum to address this problem.

When Project North Star was originally formed in 2004, the Association was
treated with considerable skepticism by the Museum. It was felt that restoring
an aircraft of the size and complexity of a North Star, to the exacting
standards of a national museum, was likely to be beyond the capabilities of a
group of volunteers and that the project would probably fail.

We will soon have completed 70,000 hours of volunteer work over a period of 14
years, restoring everything we do, to what is acknowledged to be a very high
standard. Skepticism is a thing of the past and we are now treated as a
respected and ongoing part of the Museum operations.

Respect has to be earned and comes through hard work and in the case of aircraft
restoration work, attention to detail. Our team of volunteers needs guidance
both from the Museum staff and from our own senior volunteers. One volunteer has
been particularly important in helping us the get to the position we are now in.
He has recently reached a milestone only reached on two other occasions by a
volunteer and in each case the volunteer worked in the the Science and
Technology Museum. Our volunteer has reached 10,000 volunteer hours of working
on the North Star and throughout this time he also acted as our Project Manager
for everything except the engines. Bruce Gemmill probably was not aware that he
had reached 10,000 hours when he unfortunately had to announce that he would
have to suspend his activities as a volunteer and director of PNSAC due to
family circumstances. We very much hope that he will soon be able to rejoin us
and continue to help us complete this large project. In the meantime Bruce will
continue to be our Membership Secretary. During June, Bruce's remarkable
contribution was recognized by the Museum and in a private event he was
presented with a certificate and a gift by Cedric St Amour, the Museum
Corporation Volunteer Coordinator. One of our volunteers, Robert Desjardins,
also joined Cedric on the occasion to congratulate Bruce on behalf of PNSAC.
Well done Bruce and we are all very proud of your achievement.

I have always seen our Association as having multiple roles. First and foremost
we are restoring a major artifact with our volunteers; secondly we show off our
restoration work to the public which in turn brings visitors and revenue into
the Museum and thirdly through our experienced volunteers we help to transfer
skills to younger and less experienced volunteers.

Planning our work in conjunction with the Museum management and the conservation
staff is something we must always try to jointly do better. It is often a fine
balancing act to decide when to do work on the aircraft in relation to when the
Museum is having a major event such as Canada Day. I am hoping that in future we
can regularly update our work schedule in conjunction with our project plan so
that we can more readily have the aircraft available for public viewing at the
times when the Museum would like us to do so, without having to delay our
restoration work unnecessarily.

I will conclude with an invitation to any of our members not volunteering at
present to come and join us. Although we have a definite limit on the number of
volunteers who can get their hands dirty working on restoration work, we always
welcome others who would like to participate in the running of the organization.
By becoming involved either as a restoration volunteer or an association
volunteer you will meet other like minded people and have an interesting time
talking to members of the public. You do not need to be an expert on aviation or
on the North Star to make a contribution. I can be reached by email or phone if
you would like to find out more about volunteering.

>>>>

The last NStar Chronicle was published in June 2016 and since that time we have
continued our work on the aircraft, but at a greatly reduced pace. Last summer
the Museum staged a Star Trek event in the Reserve Hanger which affected our
ability to work on the North Star. We were not asked to suspend our work until
the end of September, unlike in 2014 when preparations for the Star Wars event
were being made. This helped us retain our Restoration volunteers and recruit
two new ones.

At the end of September, Mike Irvin, the Museum's Conservation Special Projects
Manager, retired and we were asked to suspend work on the aircraft until a new
manager was appointed. In January, R\'{e}jean (Rej) Demers was appointed as the new
Special Projects Manager. Our work on the aircraft resumed on January 30th.

When Project North Star was originally formed in 2003, the Association was
treated with considerable scepticism by the Museum. It was felt that restoring
an aircraft of the size and complexity of a North Star, to the exacting
standards of a national museum, was likely to be beyond the capabilities of a
group of volunteers and that the project would probably fail.

We have now completed approximately 70,000 hours of volunteer work over a period
of 14 years, restoring everything we do to what is acknowledged to be a very
high standard. Scepticism is a thing of the past and we are now treated as a
respected and ongoing part of the Museum operations.

Respect must be earned and comes through hard work and in the case of aircraft
restoration work, attention to detail. Our team of volunteers needs guidance
both from the Museum staff and from our own senior volunteers. One volunteer has
been particularly important in helping us the get to the position we are now in.
Last year he reached a milestone only achieved on two other occasions by a
volunteer and in each case the volunteer worked at the Science and Technology
Museum. Our volunteer has worked 10,000 volunteer hours on the North Star and
for most of this time he also acted as our Project Manager for everything except
the engines. Bruce Gemmill probably was not aware that he had reached 10,000
hours when during June, his remarkable contribution was recognized by the Museum
and in a private event he was presented with a certificate and a gift by Cedric
St-Amour, the Museum Corporation Volunteer Co-ordinator.

I have always seen our Association as having multiple roles. First and foremost,
we are restoring a major artifact with our volunteers. Secondly, we show off our
restoration work to the public which in turn brings visitors and revenue into
the Museum and thirdly through our experienced volunteers we help transfer
skills to younger and less experienced volunteers.

Planning our work in conjunction with the Museum management and the Conservation
staff is something we must continuously work at. It is often a fine balancing
act to decide when to do work on the aircraft in relation to when the Museum is
having a major event such as Open Doors Ottawa and Canada Day. I am hoping that
in future we can regularly update our work schedule in conjunction with our
project plan so that we can more readily have the aircraft available for public
viewing at the times when the Museum would like us to do so, without having to
delay our restoration work unnecessarily.

During the last year, we were able to hold two of our successful Lecture Series
events and are planning to hold another one in May of this year. More details of
the next event will be available soon.

I will conclude with an invitation to any of our members not volunteering at
present to come and join us. Although we have a limit on the number of
volunteers who can get their hands dirty working on restoration, we always
welcome others who would like to join us at our events. You will meet other
likeminded people and have an interesting time talking to members of the public.
You do not need to be an expert on aviation or on the North Star to contribute.
You can contact me by email at the address at the end of this newsletter if you
would like to find out more about volunteering.

### 2017-03-15 pnsac/newsletter/article_ProgReport.tex
Draft 3--sent to Roger for review.

\title{Project Manager's Progress Report}
\subtitle{PNS 2016 Status Update}
\author{Bruce Gemmill}

@@

\begin{figure}[htbp]
   \vspace{2em}
   \centering
   \includegraphics[scale=0.5]{"PNS 007-im2".png}
   \caption*{\small \em Volunteer Peter Trobridge inspecting the air intake fairing from engine \#4 prior to beginning restoration work.}
   \label{fig:engine_no_4_2}
\end{figure}

>>>>

\title{Project Progress Report}
\subtitle{PNS 2016 Status Update}
\author{Bruce Gemmill, Peter Trowbridge, Garry Dupont}

@@

\begin{figure}[htbp]
   \vspace{2em}
   \centering
   \includegraphics[scale=0.5]{"PNS 007-im2".png}
   \caption*{\small \em Volunteer Peter Trowbridge inspecting the air intake fairing from engine \#4 prior to beginning restoration work.}
   \label{fig:engine_no_4_2}
\end{figure}
